Monte Carlo Bootstrap: current developments.

### 2019-03-02 draft.tex
updated draft. logdet+NoV working up to 12

\section{Introduction}
The Conformal Bootstrap offers a way to chart the space of acceptable CFTs.

@@

\section{LogDet implementation}
\subsection{naive approach}
Given a putative spectrum, the determinant of \ref{matrix-op} for $Nz=N_{Op}$
should get closer to zero as we approach a physically sensible set of operator
dimensions. Since this quantity could range over many orders of magnitude, we
decided to square it and then take the $\log$ of it. This proved to be
numerically troubling due to the behaviour of the logarithm around $0$  and the landscape was thus full of spurious minima out of
which the Metropolis algorithm could not get out.

@@

The next breakthrough in the routine came when we decided to start with few
($\sim 4$) operators, find the operator dimensions which minimized the 
action and then add the next operators one by one. To avoid spoiling
the previous solution, only the new operator was varied for a given number of
steps before we allowed the whole spectrum to be optimized. This gave a very
efficient routine with a large radius of convergence.

This recursive step was further improved by using
\ref{error-bound} to bound \ref{matrix-boo} as 
\begin{equation}
  | \mathcal{M}\cdot \vec{\rho}-\vec{\sigma} | < \vec{\epsilon}_{\mathrm{max}},
  \label{cross-check}
\end{equation}
where $\vec \epsilon _{\mathrm{max}}$ is just $\vec \epsilon$ under the
substitution $\mathcal{E}(z,\bar z;
\Delta^*)\to\mathcal{E}_{\mathrm{max}}(z,\bar z; \Delta^*)$.
Then, the result of each end state of the MC could be tested to see if it
satisfied Crossing Symmetry with the precision correspoinding to the $\Delta^*$
considered\footnote{$\vec \rho$ was determined by a modified $\chi^2$
minimization program. See \ref{cWLS}.}. Otherwise, it changed the seed of the pseudorandom generator and
started again the MC. The routine ended if at a given number of operators, more
than a certain number of repetitions was achieved, thus avoiding waste of CPU
time.

@@

It is an interesting fact observed in the cases studied so far, that this
constrained solution doesn't satisfy crossing with the error estimate
corresponding to the original $\Delta^*$, but once we consider the zeros in
$\vec{\rho}$, which effectively remove higher dimensional operators from the
OPE, \ref{cross-check} is satisfied.

For properly chosen
points in $z$, not all the inequalties in \ref{cross-check} will be satisfied,
one can then take this number and try to minimize it with another Monte Carlo
routine. \emph{I'm still missing consistent results for this part.}

>>>>

\section{Introduction}
The Conformal Bootstrap offers a way to chart the space of unitary and crossing
symmetric CFTs.

@@

\section{LogDet implementation}
\subsection{naive approach}
Given a putative spectrum, the determinant of \ref{matrix-op} (when
$Nz=N_{Op}+1$)
should get closer to zero as we approach a physically sensible set of operator
dimensions. Since this quantity could range over many orders of magnitude, we
decided to square it and then take the $\log$ of it. This proved to be
numerically troubling due to the behaviour of the logarithm around $0$  and the landscape was thus full of spurious minima out of
which the Metropolis algorithm could not get out.

@@

The next breakthrough in the routine came when we decided to start with few
($\sim 4$) operators, find the operator dimensions which minimized the 
action and then add the next operators one by one. To avoid spoiling
the previous solution, only the new operator was varied for a given number of
steps before we allowed the whole spectrum to be optimized. This gave a very
efficient routine with a large radius of convergence. (As of 19/02, we have
convergence for 50\% above exact dimension for $l=0,2,4,6$ and 10\% for $8\leq l
\leq 48$, using 3000 iterations for the first step and 1500 for each subsequent
one).

For unitary theories, this recursive step was further improved by using
\ref{error-bound} to bound \ref{matrix-boo} as 
\begin{equation}
  | \mathcal{M}\cdot \vec{\rho}-\vec{\sigma} | < \vec{\epsilon}_{\mathrm{max}},
  \label{cross-check}
\end{equation}
where $\vec \epsilon _{\mathrm{max}}$ is just $\vec \epsilon$ under the
substitution $\mathcal{E}(z,\bar z;
\Delta^*)\to\mathcal{E}_{\mathrm{max}}(z,\bar z; \Delta^*)$.
Then, the result of each end state of the MC could be tested to see if it
satisfied Crossing Symmetry with the precision correspoinding to the $\Delta^*$
considered\footnote{$\vec \rho$ was determined by a modified $\chi^2$
minimization program. See \ref{cWLS}.}. Otherwise, it changed the seed of the pseudorandom generator and
started again the MC. The routine ended if at a given number of operators, more
than a certain number of repetitions was achieved, thus avoiding waste of CPU
time. To avoid considering solutions with smaller $\Delta^*$ at each iteration,
we added the explicit condition of letting the number of zero entries in $\vec
\rho$ increase at most in one from run to run. This routine managed to get
relatively good results up to $l=18$, where it became increasingly difficult to
satisfy the crossing equations.

@@

\subsection{Number of Violations (NoV) minimization}
It is an interesting fact observed in the cases studied so far, that this
constrained solution doesn't satisfy crossing with the error estimate
corresponding to the original $\Delta^*$, but once we consider the zeros in
$\vec{\rho}$, which effectively remove higher dimensional operators from the
OPE, thus reducing $\Delta^*$, \ref{cross-check} is satisfied.

For properly chosen
points in $z$, not all the inequalties in \ref{cross-check} will be satisfied,
one can then take this number and try to minimize it with another Monte Carlo
routine. Concretely, this routine takes the naive $\Delta^*$ and given a certain
number of $z$-points, it counts for how many of these points are the crossing
equations are violated (NoV). It also counts how many zero coefficients are there in
the solution. Then, the optimization proceeds by changing one of the dimensions
whose corresponding OPE coefficients are not zero and then calculating the NoV
and the number of zero coefficients again. If both the NoV and the number of
zero OPE coefficients remain constant or decrease, the move is accepted,
otherwise it is rejected. The optimization terminates when NoV=0 for two
different sets of z points or when a certain number of iterations is attained.

\section{State of the art}
The best routine we have so far starts with 4 operators, uses the logdet routine
to find an approximate set of dimensions and then takes this as a starting
condition for a NoV minimization, searching to refine the solution and obtain a
higher number of non-zero OPE coefficients. To this solution a higher operator
is appended and this new spectrum is taken as the initial condition of another
logdet MC. This continues as long as the solution satisfies crossing at every
point (using the effective $\Delta^*$) and each successive step does not add
more than one extra zero OPE coefficient. I.e.
$\mathrm{zeros}(\{\Delta_l'\})\leq \mathrm{zeros}(\{\Delta_l\})+1$.
